Environmental warming increases the absolute and relative importance of high-turnover energy channels in stream food webs

### 2022-09-27 doc/ms/Junker_temp-energy-flux_ms.Rmd
ms draft full with round of comments

---
title: "Warming non-randomly speeds up food web fluxes in both an absolute and relative sense"
author:
  - James R. Junker:
      email: jrjunker@mtu.edu
      institute: [msu, glrc]
      correspondence: true
  - name: Wyatt F. Cross
    institute: msu
  - name: James M. Hood
    institute: osu
  - name: Jonathan P. Benstead
    institute: bama
  - name: Alex D. Huryn
    institute: bama
  - name: Daniel Nelson
    institute: usu
  - name: Jon S. Olafsson
    institute: veidi
  - name: Gisli M. Gislason
    institute: reyk
institute:
  - msu: Department of Ecology, Montana State University, Bozeman, MT USA
  - glrc: Current `address:` Great Lakes Research Center, Michigan Technological University, Houghton, MI USA
  - osu: Department of Evolution, Ecology, and Organismal Biology, The Ohio State University, Translational Data Analytics Institute, The Aquatic Ecology Laboratory, Columbus, OH 43212, USA
  - bama: Department of Biological Sciences, University of Alabama, Tuscaloosa, AL  35487, USA
  - usu: National Aquatic Monitoring Center, Department of Watershed Sciences, Utah State University, Logan, UT 84322, USA
  - veidi: Institute of Marine and Freshwater Fisheries, Reykjavik, Iceland
  - reyk: Univeristy of Iceland, Institute of Life and Environmental Sciences, Reykjavik, Iceland
output:
   word_document:
      reference_docx: working_docx_template.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
   pdf_document:
     keep_tex: true
   html_document: default
   pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
header-includes:
- \usepackage{lineno}
- \usepackage{amsmath}
- \usepackage{indentfirst}
- \linenumbers
indent: true
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

@@

<!--- this is a note in Rmarkdown --->

# Abstract

Warming temperatures are altering communities and trophic networks globally. While the influence of warming in food webs is often context-dependent, increasing temperatures are predicted to alter food webs and ecosystems in a number of general and fundamental ways, among them, 1) decreasing organism body size and 2) increasing metabolic rates. Both body size and metabolic rate are important drivers of many ecological patterns. Therefore, warming-induced changes have the potential to propagate through countless ecosystem processes thereby altering the distribution of food webs fluxes, food web stability, and potentially the importance of environmental and biological factors in community assembly. Here, we quantify the patterning and relative distribution of organic matter fluxes through stream food webs spanning a natural ~25$^\circ$C temperature gradient. We then related patterns in fluxes to species and community traits (body size, *M* and biomass turnover, *P:B*) within and across ecosystems. We predicted 1) fluxes would skew towards smaller *M* and higher *P:B* both across and within ecosystems and 2) fluxes would be more skewed at higher temperatures. Across the temperature gradient, communities in warmer streams were composed of smaller-bodied and higher-turnover populations on average. Additionally, material fluxes of warmer streams were increasingly skewed towards high-turnover populations *within* the community, showing higher temperatures restructured organic matter fluxes in both an absolute and relative sense. Lastly, in warmer ecosystems the relative distribution of organic matter fluxes appears to be increasingly 'non-random', evidence of potentially stronger selection for these traits with increasing temperature. The emerging picture is a warmer world that is both smaller and faster. Our study lends further support to this pattern and suggests that temperature will become an increasingly important environmental filter on warming communities in the future.
\newpage

@@

<!--- warming and effects in food webs --->
<!--- Speeding up of food web: Species levels--->
Increasing global temperatures are altering the provisioning and maintenance of ecosystem services by modifying the network of interactions among species that underpin ecosystem functions [@brose2012; @woodward2010; @deruiter1995; @thompson2012]. Across global climate gradients, temperature mediates the dynamics and stability of food webs [@baiser2019] by altering the acquisition and allocation of resources among species [@zhang2017] and ultimately the magnitudes and relative distribution of biomass and fluxes among species and trophic levels [@may1972; @mccann1998; @barnes2018; @gibert2019a]. Temperature's effects permeate across levels of biological organization from its control on individual metabolic rates [@gillooly2001; @brown2004] and biological activity (e.g., attack rate, handling time, growth rates, etc.; @dell2014, as an environmental niche filter of community assembly and indirectly through the distribution of species traits, such as body size, within and across species [@bideault2019; @nelson2017].

<!--- Connect traits to the 'speed' of ecosystem energy fluxes--->
In fact, reduced organismal body size with warming is considered a 'universal' response to climate change [@atkinson1994; @gardner2011; @daufresne2009]. Body size reductions arise from multiple mechanisms acting at different levels of organization, such as the increased relative abundance of smaller species within warmer communities [@bergmann1848], or smaller individuals in warmer populations [@james1970], or through reduced body size of warmer individuals [@atkinson1994]. Temperature-size relationships (TSR) are likely to influence the distribution of body sizes both across and within ecosystems, such that warmer communities should contain smaller species and populations, on average, and smaller organisms should be more dominant within a community. These changes can have important implications for ecosystems pattern and processes because body size is a strong determinant of an organism's biology [@peters1983] influencing, life history patterns [@zeuss2017; @altermatt2010; @nelson2020b], and developmental [@angilletta2004] and metabolic rates [@brown2004; @gillooly2001].

<!--- Write the main goals of the paragraph--->
Organismal metabolic rate, in addition to being controlled by body size, is strongly influenced by temperature [@gillooly2001]. Therefore, temperature may alter metabolic rate indirectly through reductions in body size, as well as directly through its effects on subcellular kinetics [@bideault2019; @osmond2017]. The interactive effect of these processes has been shown to modulate ecosystem patterns through effects on population abundance [@bernhardt2018], consumer-resource interactions [@bideault2019], and food web structure [@gibert2019a]. Importantly, the kinetic effects of temperatures are tied to many biological processes [@dell2014], among them growth rate [@gillooly2001], developmental rate [@zuo2012; @nelson2020b], voltinism [@zeuss2017], and population turnover rate [@huryn2007; @brown2004]. As such, the emerging picture suggests a smaller, faster world in which warming "speeds up" ecosystem processes through the compounded effects of smaller body size and higher turnover rates.*possibly simplify the paragraph and combined with above*

Yet, considerable uncertainty remains in the predicted short- and long-term effects of climate warming, especially at higher levels of organization such as communities and food webs [@woodward2010; @walther2002]. For example, species interactions in food webs may modulate species' TSRs [@delong2014], leading to varied responses among species and trophic levels [@ohlberger2013]. Additionally, warming-induced body size reductions may act as a stabilizing process in some consumer-resource interactions through reduced consumer:resource biomass ratios [@osmond2017] and reductions in consumer growth rates relative to resource supply [@kozlowski2004; @delong2012a; @mccann2011; @greyson-gaito2022]. 
Therefore, food web processes may impart varying selective pressures on the distribution of body sizes and population turnover within communities. If strong, these pressures may leave distinct imprints on community structure and the distribution of energy fluxes within food webs with potential implications for the stability and functioning of ecosystems on a warming planet [@petchey1999; @fussmann2014].

Here, we measured the patterning and distribution of energy fluxes within invertebrate food webs across a natural stream temperature gradient (\~5--28$^\circ$C). Previous research in these streams has shown a strong positive effect of temperature on primary production both among streams [@demars2011; @padfield2017] and within streams seasonally
[@ogorman2012; @hood2018]. Consumers rely largely on autochthonous resources [@ogorman2012; @nelson2020] and therefore the dynamics of primary production have a strong control on consumer energy demand [@junker2020]. As such, we predicted annual energy fluxes of consumers to scale with among-stream patterns in resource availability and consumer energy demand and increase with temperature across streams. We hypothesized temperature to be a strong environmental filter of community assembly and the distribution of energy and organic matter fluxes, especially at the extremes. From this we made a number of predictions regarding the patterning of energy fluxes within and across communities of different temperature, with particular focus on two species traits: body size and population biomass turnover (production:biomass ratios, *P:B*). First, we predict warmer temperatures should favor smaller and faster turnover species leading to a decrease in body sizes and increase in *P:B* on average across communities. Second, within communities energy fluxes should be dominated by smaller, higher turnover species at warmer temperatures. Third, we hypothesized that the selective pressure from environmental temperature on body sizes and population turnover would be greatest at temperature extremes, therefore, the patterning of energy fluxes will be 'non-random' at coldest and warmest temperatures.

@@

Daily secondary production of invertebrate taxa was calculated using the instantaneous growth rate method [IGR, @benke2017]. Growth rates were determined using taxon appropriate
approaches described in [@junker2020]. Briefly, growth rates of common taxa (e.g., Chironomidae spp., *Radix balthica*, etc.) were determined using *in situ* chambers [@huryn1986].
Multiple individuals (*n* = 5--15) within small size categories (\~1 mm length range) were photographed next to a field micrometer, placed into the stream within pre-conditioned
chambers for 7--15 days, after which they were again photographed. Individual lengths were measured from field pictures using image analysis software [@schindelin2012], and body
lengths were converted to mass (mg ash-free dry mass [AFDM]) using published length-mass regressions [@benke1999; @ogorman2012; @hannesdottir2013]. Growth rates (*g*, d^-1^) were
calculated by the changes in mean body size (*W*) over a given time interval (*t*) with the following equation:

$$g = log_e ( W_{t+\Delta t} / W_t) / \Delta t$$

Variability in growth rates was estimated by bootstrapping through repeated resampling of individual lengths with replacement (*n* = 1000). For taxa which exhibit synchronous growth
and development (e.g., Simuliidae spp., some Chironomidae spp., etc.), we examined temporal changes in length-frequency distributions and calculated growth rates and uncertainty
using a bootstrap technique similar to that described in Benke and Huryn [-@benke2017]. Individual Lengths were converted to mg AFDM using published length-mass regression cited
above and size-frequency histograms were visually inspected for directional changes in body size through time. For each date, size-frequency distributions were resampled with
replacement and growth rates estimated from equation 1. We prevented the calculation of negative growth rates by requiring $W_{t + \Delta t}$ \> $W_t$. To estimate
growth rates of taxa for which growth could not be estimated empirically, we developed stream-specific growth rate models by constructing multivariate linear regressions of
empirical growth data against body size and temperature within each stream. To estimate uncertainty in production of each taxon, we used a bootstrapping technique that resampled
measured growth rates, in addition to abundance and size distributions from individual samples. For each iteration, size-specific growth rates were multiplied by mean
interval biomass for each size class and the number of days between sample dates to estimate size class-specific production. For each interval, size classes were summed for each
taxon to calculate total population-level production. Intervals were summed to estimate annual secondary production.

## Diet analysis

From modeled diet compositions, we estimated trophic redundancy within and across stream food webs by calculating proportional similarities (*PS*; @whittaker1952 among modeled diet estimates. Proportional similarities were calculated as:

@@

Consumption fluxes (g m^-2^ t^-1^) were calculated using the trophic basis of production method [TBP, @benke1980]. Taxon-specific secondary production estimates were combined with
diet proportions, diet-specific assimilation efficiencies, *AE~i~*, and assumed net production efficiencies, *NPE*, to estimate consumption of organic matter. For each food
category, *i*, diet proportions were multiplied by the gross growth efficiency ($GGE_{i} = AE_{i} * NPE$) to calculate the relative production attributable to each food
category. The relative production from each food type was then multiplied by the interval-level production and finally divided by $GGE_{i}$ to estimate consumption of organic
matter from each food category by a consumer [@benke1980]. Consumption was calculated for each taxon across sampling intervals (typically \~1 month). Total interval consumption was
calculated by summing across all taxa, while annual consumption was calculated by summing across all taxa and intervals. Variability in consumption estimates was estimated through
a Monte Carlo approach, wherein bootstrapped vectors of secondary production for each taxon (see *Secondary production* methods above) were resampled and consumption estimated with
the TBP method using modeled diet proportions (see *Diet analysis* above), diet-specific assimilation efficiencies, and net production efficiency. Variability in *AE~i~* was
incorporated by resampling values from beta distributions fit to median and 2.5% and 97.5% percentiles for each diet item: diatoms = 0.30 (95% percentile interval (PI): 0.24-0.36),
filamentous and green algae = 0.30 (95% PI: 0.24-0.36), cyanobacteria = 0.10 (95% PI: 0.08-0.12), amorphous detritus = 0.10 (95% PI: 0.08-0.12), vascular and non-vascular plants
(bryophytes) = 0.1 (95% PI: 0.08-0.12), and animal material = 0.7 (95% PI: 0.56-0.84)[@welch1968; @cross2007; @cross2011; @benke1980; @benke1997]. Variability in *NPE* was
incorporated by resampling values from an assumed beta distribution with median *NPE* = 0.45 (95% PI = 0.4-0.5). Beta distributions were fit using the 'get.beta.par()' function
within the *rriskDistributions* package [@belgorodski2017].

@@

To visualize and quantify how evenly OM fluxes were distributed among consumers within a stream, we constructed Lorenz curves [@lorenz1905] on ordered relative consumption fluxes, such that in a community with $S$ species and the relative consumption of species *i*, $p_i$, is ordered $p_1 \leq p_2 \leq ... p_S$. The Lorenz curve plots how a value, in this case OM flux, accumulates with increasing cumulative proportion of species. In a community with perfectly equal distribution of OM consumption among species, the Lorenz curve is simply a straight diagonal line. Deviation from perfect equality was calculated as the Gini coefficient [@gini1921], normalized for differences in $S$ among streams, $G^*$ [@solomon1975; @chao2019]:

@@

We predicted that warming would favor species with smaller body size and higher population turnover and therefore OM fluxes would be skewed towards small body size (*M*) and higher *P:B* across and within communities. Across communities, we assessed the relationships between mean annual temperature ($^\circ$C) and mean population *M* and *P:B* of each community with bootstrapped linear regressions. Here, `r temperature_stats[['n_boot']]` values of population *M* or *P:B* were resampled with replacement for each species within each stream. The mean of all populations within a stream was calculated and a linear model was fit between *log~e~*-transformed *M* or *P:B* and mean annual temperature. Response variables *M* and *P:B* were transformed to meet the assumption of normally distributed residual variation.

To quantify the distribution of energy fluxes _**within**_ a community, we assessed if relative OM flux was skewed towards populations with lower or higher relative *M* or  *P:B*. To do this, we ordered species based on within-stream ranking of annual population traits (i.e, $M$, $P:B$) and then calculated a measure of skewness, $Sk_{flux}$, based on quartiles of the distribution of OM fluxes in relation to species traits as:

@@

To assess the strength of environmental filtering of species' traits within a community, we quantified how likely it was to observe such a skewed distribution in OM fluxes by random chance. Here, we predicted that communities with strong filtering for traits would exhibit highly skewed OM distributions that would be unlikely due to chance (i.e., 'non-random ordering'). In contrast, weak environmental filtering would produce OM fluxes with skew values that are likely due to random chance ('random ordering'). To accomplish this, we first had to account for statistical constraints that restrict the range of possible outcomes [\ i.e., feasible set, @haegeman2008; @diaz2021a], given the number of species and the relative distribution of OM fluxes within a community. The number of unique orderings of species increases to computationally intractable numbers very quickly (e.g., $S!$, 10 species ~3.6e^6^ unique orderings). Therefore, we chose to permute a random subset of each stream community's feasible set by randomly ordering species and calculating the skewness in the cumulative distribution of annual OM fluxes 100,000 times in each stream. The number of random orderings was chosen as a balance between characterizing the distribution of skewness values within each feasible set and computational and time constrains. This permuted set allowed us to calculate the probability of observing the empirical skewness, $Sk_{flux}$, in each stream compared to a random ordering given the distribution of relative OM flux. To assess the relationship between the probability of 'non-random ordering' and temperature, we again used beta regression on estimates of the probability of seeing the empirical $Sk_{flux}$ value or one more extreme based on random ordering of species in the permuted feasible set. Here, values can take the range between 0 and 1 so no transformation was necessary. As necessary, when we observed many values near 0 and 1, we implemented a zero-one inflated beta regressions in brms [@burkner2017] to test if the probability of observing the empirical $Sk_{flux}$ was associated with mean annual temperature across streams. We predicted increasingly 'non-random' distributions in OM fluxes with increasing temperature.

@@

## Community OM fluxes

Annual community energy demand mirrored patterns of secondary production previously reported [@junker2020]. Community energy demand varied ~`r round((max_demand$flux_mg_m_y_mean/min_demand$flux_mg_m_y_mean),1)`-fold across streams (`r round((min_demand$flux_mg_m_y_mean/1000), 3)`; `r round((min_demand$flux_mg_m_y_quant2.5/1000),3)` -- `r round((min_demand$flux_mg_m_y_quant97.5/1000),3)` to `r round((max_demand$flux_mg_m_y_mean/1000),1)`; `r round((max_demand$flux_mg_m_y_quant2.5/1000),1)` -- `r round((max_demand$flux_mg_m_y_quant97.5/1000),1)` g AFDM $m^{-2} y^{-1}$ mean; 95% percentile interval (PI)) and was positively related to temperature. Total OM flux through the consumer community was strongly associated with community energy demand and varied from `r round((min_flux$flux_mg_m_y_mean/1000),1)` (`r round((min_flux$flux_mg_m_y_quant2.5/1000),1)`--`r round((min_flux$flux_mg_m_y_quant97.5/1000),1)`, 95% PI) to `r round((max_flux$flux_mg_m_y_mean/1000),1)` (`r round((max_flux$flux_mg_m_y_quant2.5/1000),1)`--`r round((max_flux$flux_mg_m_y_quant97.5/1000),1)`, 95% PI) g AFDM $m^{-2} y^{-1}$.

@@

## Evenness of OM fluxes within streams

>>>>

---
title: "Environmental warming increases the absolute and relative importance of high-turnover energy channels in stream food webs"
author:
  - James R. Junker:
      email: jrjunker@mtu.edu
      institute: [msu, glrc]
      correspondence: true
  - name: Wyatt F. Cross
    institute: msu
  - name: James M. Hood
    institute: osu
  - name: Jonathan P. Benstead
    institute: bama
  - name: Alex D. Huryn
    institute: bama
  - name: Daniel Nelson
    institute: usu
  - name: Jón S. Ólafsson
    institute: veidi
  - name: Gísli M. Gíslason
    institute: reyk
institute:
  - msu: Department of Ecology, Montana State University, Bozeman, MT 59717, USA
  - glrc: Current `address:` Great Lakes Research Center, Michigan Technological University, Houghton, MI 49931, USA
  - osu: Department of Evolution, Ecology, and Organismal Biology, The Ohio State University, Translational Data Analytics Institute, The Aquatic Ecology Laboratory, Columbus, OH 43212, USA
  - bama: Department of Biological Sciences, University of Alabama, Tuscaloosa, AL  35487, USA
  - usu: National Aquatic Monitoring Center, Department of Watershed Sciences, Utah State University, Logan, UT 84322, USA
  - veidi: Institute of Marine and Freshwater Fisheries, Reykjavik, Iceland
  - reyk: Univeristy of Iceland, Institute of Life and Environmental Sciences, Reykjavik, Iceland
output:
   word_document:
      reference_docx: working_docx_template.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
   pdf_document:
     keep_tex: true
   html_document: default
   pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
header-includes:
- \usepackage{lineno}
- \usepackage{amsmath}
- \usepackage{indentfirst}
- \linenumbers
indent: true
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

@@

<!--- this is a note in Rmarkdown --->
# Abstract

Warming temperatures are altering communities and trophic networks globally. While the influence of warming on food webs is often context-dependent, increasing temperatures are predicted to influence communities in two fundamental ways: 1) by reducing average body size and 2) by increasing individual metabolic rates. Both body size and metabolic rate are important drivers of many ecological patterns. Therefore, warming-induced changes have the potential to propagate through countless ecosystem processes thereby altering the distribution of food web fluxes, food web stability, and potentially the importance of environmental and biological factors in community assembly. Here, we quantify the patterning and relative distribution of organic matter fluxes through stream food webs spanning a natural ~25$^\circ$C temperature gradient. We then related patterns in fluxes to species and community traits (body size, *M* and biomass turnover, *P:B*) within and across ecosystems. We predicted 1) warmer communities would skew towards smaller *M* and higher *P:B* across ecosystems, 2) organic matter fluxes within communities would increasingly skew towards smaller, higher *P:B* populations at higher temperatures. Across the temperature gradient, communities in warmer streams were composed of smaller-bodied and higher-turnover populations on average. Additionally, material fluxes of warmer streams were increasingly skewed towards high-turnover populations *within* the community, showing higher temperatures restructured organic matter fluxes in both an absolute and relative sense. With warming, the relative distribution of organic matter fluxes also appeared to be increasingly 'non-random', suggesting the potential for stronger selection for these traits with increasing temperature. The emerging picture is a warmer world that is both smaller and faster. Our study provides clear empirical support for this pattern in natural communities and suggests that warming has the potential to shape pathways of energy and material flows in food webs in ways that may be predictable.

@@

<!--- warming and effects in food webs --->
<!--- Speeding up of food web: Species levels--->
Increasing global temperatures can influence the provision and maintenance of ecosystem services by modifying the network of species interactions that underpin ecosystem functions [@brose2012; @woodward2010; @deruiter1995; @thompson2012]. The effects of warming permeate across levels of biological organization from its control on individual metabolic rates [@gillooly2001; @brown2004] and biological activity (e.g., attack rate, handling time, growth rates, etc.; @dell2014, to broad-scale shifts in community assembly and structure [@saito2021; @gibert2019a; @nelson2017a]. Across global climate gradients, these temperature-induced changes have the potential to mediate food web stability [@baiser2019] by altering the acquisition and allocation of resources among species [@zhang2017] and the magnitudes and relative distribution of energy fluxes among species and trophic levels [@may1972; @mccann1998; @barnes2018].

Understanding how warming alters energy fluxes in food webs requires information about how temperature modifies connections between ecosystem structure (i.e. biodiversity) and function (BEF). The extent to which temperature modifies the relative importance of community assembly processes will mediate changes to ecosystem structure and function with warming within and among ecosystems. These processes are varied--from strictly deterministic [\ e.g., environmental/niche filtering, @whittaker1962],  to strictly stochastic [\ e.g., neutral theory, @hubbell2001]--and they operate across multiple scales. As such, there are a multitude of possible relationships between warming and species' abundance and trait distributions. For example, strong environmental filtering can be expected to mediate the number of species in a community and is likely to skew trait distributions, to some extent, in natural communities [@therriault1999]. However, the relative distribution of traits can be modified by a number of additional processes beyond environmental filtering that alter species relative abundance distributions [\ e.g., species interactions @therriault1999 and demographic stochasticity @hubbell2001]. These additional processes can exaggerate or counter any skew in trait distribution imparted through environmental filtering and may therefore be important for understanding the warming response in species and communities and the absolute and relative distribution of organic matter fluxes across gradients in species traits.

Although temperature appears to have minimal control on species richness across taxonomic groups [e.g., @bastazini2021], warming may have profound effects on the dominance structure (i.e. evenness) of ecological communities, often by reducing community evenness and favoring a few well-adapted species [@hillebrand2008]. The implications of these changes on the magnitude and pathways of energy flux in food webs should depend on how warming influences functional trait distributions, as well as how these functional traits translate to energy demands [@loreau2001; @norberg2001; @norberg2004].

<!--- Connect traits to the 'speed' of ecosystem energy fluxes--->
Body size is a fundamental trait that is strongly influenced by temperature and has great potential to influence energy flux [@atkinson1994; @gardner2011; @daufresne2009]. Body size reductions arise from multiple mechanisms acting at different levels of organization, such as the increased relative abundance of smaller species within warmer communities [@bergmann1848], or smaller individuals in warmer populations [@james1970], or through reduced body size of warmer individuals [@atkinson1994]. Temperature-size relationships (TSR) are likely to influence the distribution of body sizes both across and within ecosystems, such that warmer communities should contain smaller species and populations, on average, and smaller organisms should be more dominant within a community. These changes can have important implications for ecosystems pattern and processes that control energy and material fluxes in ecosystems because body size is a strong determinant of an organism's biology [@peters1983] influencing, life history patterns [@zeuss2017; @altermatt2010; @nelson2020b], and developmental [@angilletta2004] and metabolic rates [@brown2004; @gillooly2001].

<!--- Write the main goals of the paragraph--->
Metabolic rate, in addition to being controlled by body size, is a second key trait that is influence by temperature and tied to variation in energy flux through food webs [@gillooly2001; ]. Warming can influence metabolic indirectly through reductions in body size, as well as directly through its effects on subcellular kinetics [@bideault2019; @osmond2017]. These processes has been shown to modify ecosystem patterns through changes in population abundance [@bernhardt2018], consumer-resource interactions [@bideault2019], and food web structure [@gibert2019a]. Importantly, the kinetic effects of temperature are tied to many biological processes [@dell2014], among them growth rate [@gillooly2001], developmental rate [@zuo2012; @nelson2020b], voltinism [@zeuss2017], and population turnover rate [@huryn2007; @brown2004]. Taken together, these changes suggest a smaller, faster world in which warming "speeds up" ecosystem processes through the compounded effects of smaller body size and higher turnover rates, with potentially lasting imprints on food web structure and energy fluxes [@gibert2019a].

Here, we measured the patterning and distribution of organic matter fluxes within invertebrate food webs across a natural stream temperature gradient (\~5--28$^\circ$C). Previous research in these streams has shown a strong positive effect of temperature on primary production both among streams [@demars2011; @padfield2017] and within streams seasonally [@ogorman2012; @hood2018]. Consumers rely largely on autochthonous resources [@ogorman2012; @nelson2020] and therefore the dynamics of primary production have a strong control on consumer energy demand [@junker2020]. As such, we predicted annual organic matter fluxes to consumers would increase with temperature across streams mirroring among-stream patterns in resource availability and consumer energy demand. We hypothesized temperature to be a strong environmental filter of community assembly and the distribution of energy and organic matter fluxes. From this we made a number of predictions regarding the patterning of organic matter fluxes within and across communities of different temperature, with particular focus on two species traits; body size and population biomass turnover (production:biomass ratios, *P:B*). First, we predict warmer temperatures should favor smaller and faster turnover species leading to a decrease in body sizes and increase in *P:B* on average across communities. Second, within communities organic matter fluxes should be increasingly dominated by smaller, higher turnover species at warmer temperatures. Lastly, we quantify the probability that patterns of organic matter fluxes among populations arise from chance versus the potential for temperature to structure the relative distribution of material fluxes along species trait axes.

@@

Daily secondary production of invertebrate taxa was calculated using the instantaneous growth rate method [IGR, @benke2017]. Growth rates were determined using taxon appropriate
approaches described in [@junker2020]. Briefly, growth rates of common taxa (e.g., Chironomidae spp., *Radix balthica*, etc.) were determined using *in situ* chambers [@huryn1986].
Multiple individuals (*n* = 5--15) within small size categories (\~1 mm length range) were photographed next to a field micrometer, placed into the stream within pre-conditioned
chambers for 7--15 days, after which they were again photographed. Individual lengths were measured from field pictures using image analysis software [@schindelin2012], and body
lengths were converted to mass (mg ash-free dry mass [AFDM]) using published length-mass regressions [@benke1999; @ogorman2012; @hannesdottir2013]. Growth rates (*g*, d^-1^) were
calculated by the changes in mean body size (*M*) over a given time interval (*t*) with the following equation:

$$g = log_e ( M_{t+\Delta t} / M_t) / \Delta t$$

Variability in growth rates was estimated by bootstrapping through repeated resampling of individual lengths with replacement (*n* = 1000). For taxa which exhibit synchronous growth
and development (e.g., Simuliidae spp., some Chironomidae spp., etc.), we examined temporal changes in length-frequency distributions and calculated growth rates and uncertainty
using a bootstrap technique similar to that described in Benke and Huryn [-@benke2017]. Individual Lengths were converted to mg AFDM using published length-mass regression cited
above and size-frequency histograms were visually inspected for directional changes in body size through time. For each date, size-frequency distributions were resampled with
replacement and growth rates estimated from equation 1. We prevented the calculation of negative growth rates by requiring $M_{t + \Delta t}$ \> $M_t$. If this condition was not met after 1e^4^ resamplings, a minimum growth rate of 0.001 was set. To estimate
growth rates of taxa for which growth could not be estimated empirically, we developed stream-specific growth rate models by constructing multivariate linear regressions of
empirical growth data against body size and temperature within each stream. To estimate uncertainty in production of each taxon, we used a bootstrapping technique that resampled
measured growth rates, in addition to abundance and size distributions from individual samples. For each iteration, size-specific growth rates were multiplied by mean
interval biomass for each size class and the number of days between sample dates to estimate size class-specific production. For each interval, size classes were summed for each
taxon to calculate total population-level production. Intervals were summed to estimate annual secondary production.

@@

Organic matter fluxes through the community (g m^-2^ t^-1^) were calculated using the trophic basis of production method [TBP, @benke1980]. Taxon-specific secondary production estimates were combined with
diet proportions (see Supplemental Materials), diet-specific assimilation efficiencies, *AE~i~*, and assumed net production efficiencies, *NPE*, to estimate consumption of organic matter. For each food
category, *i*, diet proportions were multiplied by the gross growth efficiency ($GGE_{i} = AE_{i} * NPE$) to calculate the relative production attributable to each food
category. The relative production from each food type was then multiplied by the interval-level production and finally divided by $GGE_{i}$ to estimate consumption of organic
matter from each food category by a consumer [@benke1980]. Consumption was calculated for each taxon across sampling intervals (typically \~1 month). Total interval consumption was
calculated by summing across all taxa, while annual consumption was calculated by summing across all taxa and intervals. Variability in consumption estimates was estimated through
a Monte Carlo approach, wherein bootstrapped vectors of secondary production for each taxon (see *Secondary production* methods above) were resampled and consumption estimated with
the TBP method using modeled diet proportions (see *Diet analysis* in Supplemental Materials), diet-specific assimilation efficiencies, and net production efficiency. Variability in *AE~i~* was
incorporated by resampling values from beta distributions fit to median and 2.5% and 97.5% percentiles for each diet item: diatoms = 0.30 (95% percentile interval (PI): 0.24-0.36),
filamentous and green algae = 0.30 (95% PI: 0.24-0.36), cyanobacteria = 0.10 (95% PI: 0.08-0.12), amorphous detritus = 0.10 (95% PI: 0.08-0.12), vascular and non-vascular plants
(bryophytes) = 0.1 (95% PI: 0.08-0.12), and animal material = 0.7 (95% PI: 0.56-0.84)[@welch1968; @cross2007; @cross2011; @benke1980; @benke1997]. Variability in *NPE* was
incorporated by resampling values from an assumed beta distribution with median *NPE* = 0.45 (95% PI = 0.4-0.5). Beta distributions were fit using the 'get.beta.par()' function
within the *rriskDistributions* package [@belgorodski2017].

@@

To visualize and quantify how evenly OM fluxes were distributed among consumers within a stream, we constructed Lorenz curves [@lorenz1905] on rank ordered OM fluxes, such that in a community with $S$ species and the relative OM flux of species *i*, $p_i$, is ordered $p_1 \leq p_2 \leq ... p_S$. The Lorenz curve plots how a value, in this case OM flux, accumulates with increasing cumulative proportion of species. In a community with perfectly equal distribution of OM flux among species, the Lorenz curve is simply a straight diagonal line. Deviation from perfect equality was calculated as the Gini coefficient [@gini1921], normalized for differences in $S$ among streams, $G^*$ [@solomon1975; @chao2019]:

@@

We predicted that warming would favor species with smaller body size and higher population turnover and therefore OM fluxes would be skewed towards small body size (*M*) and higher *P:B* across and within communities. To assess the potential effects of environmental filtered across communities, we explored the relationships between mean annual temperature ($^\circ$C) and mean population *M* and *P:B* of each community with bootstrapped linear regressions. Here, `r temperature_stats[['n_boot']]` values of population *M* or *P:B* were resampled with replacement for each species within each stream. The mean of all populations within a stream was calculated and a linear model was fit between *log~e~*-transformed *M* or *P:B* and mean annual temperature. Response variables *M* and *P:B* were transformed to meet the assumption of normally distributed residual variation.

To quantify how the distribution of organic matter fluxes were modified _**within**_ a community and how this modification was related to temperature, we assessed the extent to which the relative OM fluxes among consumers within a community were skewed towards populations with lower or higher relative *M* or *P:B*. To do this, we ordered species based on within-stream ranking of annual population traits (i.e., $M$, $P:B$) and then calculated a measure of skewness, $Sk_{flux}$, based on quartiles of the distribution of OM fluxes in relation to species traits as:

@@

To assess how temperature may structure the skew of OM among populations within a community, we quantified how likely it was to observe such a skewed distribution in OM fluxes by random chance. Skewness in OM fluxes may arise through both ecological and random processes. Further, the feasible range of skewness values a community can take is inherently tied to the evenness of OM fluxes within a community. This dependence on evenness can make it difficult to determine the importance of random versus ecological processes on the distribution of OM fluxes in the community through comparison of raw skewness measures alone. Here, we predicted that species' $M$ and $P:B$ would be increasingly important traits structuring OM fluxes within a community and therefore warmer streams would exhibit highly skewed OM distributions that would be unlikely due to chance (i.e., 'non-random ordering'). In contrast, cooler streams would exhibit OM fluxes with skew values that are more likely due to random chance ('random ordering'), regardless of raw skewnness values, suggesting other traits govern the distribution of OM fluxes within communities. To accomplish this, we first had to account for statistical constraints that restrict the range of possible outcomes [\ i.e., feasible set, @haegeman2008; @diaz2021a], given the number of species and the relative distribution of OM fluxes within a community. The number of unique orderings of species increases to computationally intractable numbers very quickly (e.g., $S!$, 10 species ~3.6e^6^ unique orderings). Therefore, we chose to permute a random subset of each stream community's feasible set by randomly ordering species and calculating the skewness in the cumulative distribution of annual OM fluxes 100,000 times in each stream. The number of random orderings was chosen as a balance between characterizing the distribution of skewness values within each feasible set and computational and time constrains. This permuted set allowed us to calculate the probability of observing the empirical skewness, $Sk_{flux}$, in each stream compared to a random ordering given the distribution of relative OM flux.

@@

## Community organic matter fluxes

Annual community OM fluxes mirrored patterns of secondary production reported previously [@junker2020]. Organic matter flux to invertebrates varied ~`r round((max_flux$flux_mg_m_y_mean/min_flux$flux_mg_m_y_mean),0)`-fold among streams (mean [95% percentile interval (PI)]: `r round((min_flux$flux_mg_m_y_mean/1000), 1)`[`r round((min_flux$flux_mg_m_y_quant2.5/1000),1)` -- `r round((min_flux$flux_mg_m_y_quant97.5/1000),1)` to `r round((max_flux$flux_mg_m_y_mean/1000),1)` [`r round((max_flux$flux_mg_m_y_quant2.5/1000),1)` -- `r round((max_flux$flux_mg_m_y_quant97.5/1000),1)`] g AFDM $m^{-2} y^{-1}$) and was positively related to temperature (Figure 1A).

@@

## Evenness of organic matter fluxes
